Leaf traits drive differences in biomass partitioning among major plant functional types

### 2014-12-15 manuscript.Rnw
Added Methods S1, explaining allocation vs. partitioning

\affil[1]{1 Hawkesbury Insitute for the Environment, University of Western Sydney, Hawkesbury Campus, First Floor, Building L9, Locked Bag 1797, Penrith NSW 2751, Australia}
\affil[2]{Biological Sciences, Macquarie University NSW 2109, Australia}
\affil[3]{See Supporting Information}

@@

\textbf{
The partitioning of biomass into short-lived leaves and long-lived stems is one of the most uncertain  and influential components of dynamic global vegetation models (DGVMs), affecting the lifetime of carbon in ecosystems \cite{friend_carbon_2014}, predictions of global carbon storage \cite{ise_comparison_2010}, and the CO\textsubscript{2} response of net primary production \cite{dekauwe2014}. Although DGVMs typically assume that the major woody plant functional types (PFTs) differ in biomass partitioning, empirical data have been lacking. Using a new database on biomass partitioning from over 175 studies, we here show a strong effect of PFT on biomass partitioning. Contrary to prior expectation, we show that major woody PFTs do not differ in the amount of leaf area supported, but rather the amount of leaf mass per unit above ground biomass. Moreover, we find that these differences relate to systematic shifts in leaf mass per area, a leading trait within the leaf economics spectrum. Combined, these results uncover fundamental rules in how plants are constructed and allow for systematic benchmarking of biomass partitioning routines in current DGVMs.
}

@@

\item DGVMs are used to predict the trajectory of the global vegetation carbon sink in a changing climate. Biomass partitioning (related to, but not to be confused with allocation of net primary production) is important because short-lived leaves re-enter the decomposition cycle quickly, releasing much of its carbon back to the atmosphere. Long-lived stems act as a very long carbon storage term (Luyssaert). As a result, partitioning affects the overall lifetime of carbon in the ecosystem (Friend). 

@@

\subsection*{Predicting basal diameter}
\label{sec:predas}

>>>>

\affil[1]{1 Hawkesbury Insitute for the Environment, University of Western Sydney, Hawkesbury Campus, First Floor, Building L9, Locked Bag 1797, Penrith, NSW 2751, Australia}
\affil[2]{Biological Sciences, Macquarie University, North Ryde, NSW 2109, Australia}
\affil[3]{See Supporting Information}

@@

\textbf{
The partitioning of biomass into short-lived leaves and long-lived stems is one of the most uncertain  and influential components of  global vegetation models (GVMs), affecting the lifetime of carbon in ecosystems \cite{friend_carbon_2014}, predictions of global carbon storage \cite{ise_comparison_2010}, and the CO\textsubscript{2} response of net primary production \cite{dekauwe2014}. Although GVMs typically assume that the major woody plant functional types (PFTs) differ in biomass partitioning, empirical data have been lacking. Using a new global database on biomass partitioning from over 175 studies, we here show a strong effect of PFT on partitioning between stems and leaves. Contrary to prior expectation, we show that major woody PFTs do not differ in the amount of leaf area supported, but rather the amount of leaf mass per unit above ground biomass. Moreover, we find that these differences relate to systematic shifts in leaf mass per area, a leading trait within the leaf economics spectrum. Combined, these results uncover fundamental rules in how plants are constructed and allow for systematic benchmarking of biomass partitioning routines in current GVMs.
}

@@

\item GVMs (including dynamic GVMs, which model the distribution of vegetation types, and those who use a prescribed vegetation distributions) are used to predict the trajectory of the global vegetation carbon sink in a changing climate. Biomass partitioning (related to, but not to be confused with allocation of net primary production) is important because short-lived leaves re-enter the decomposition cycle quickly, releasing much of its carbon back to the atmosphere. Long-lived stems act as a very long carbon storage term (Luyssaert). As a result, partitioning affects the overall lifetime of carbon in the ecosystem, which has been shown to be an important determinant of overall carbon storage projections(Friend). 

@@

\subsection*{Methods S1 - Biomass partitioning terminology}
\label{sec:partterm}

The terms 'partitioning' and 'allocation' have been used in various ways, confusing comparisons between studies \cite{litton_carbon_2007}. Here, we define biomass partitioning as the actual distribution of biomass between compartments (e.g., leaves vs. stems), and biomass allocation as the proportion of net primary production (NPP) that is allocated to some compartment. The two processes are different because of continuous turnover of biomass, which differs strongly between compartments. We may write (cf. McMurtrie and Wolf 1983),

$$
\frac{dM_i}{dt} = a_i NPP - \tau_i M_i
$$

where $M_i$ is the biomass in some compartment (leaves, stems or roots), $a_i$ the annual allocation of NPP to that compartment, and $\tau_i$ the annual turnover of compartment $i$. It is thus easy to see that partitioning ($M_i/M_T$, where $M_T$ is total biomass) can be different from allocation ($a_i$) because turnover ($\tau_i$) differs strongly between compartments.

\subsection*{Methods S2 - Predicting basal diameter}
\label{sec:predas}
